---
title: "final year projects"
author: "karthi"
date: "2024-12-10"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
```

#import the Dataset
```{r}
project=read.csv(("C:/Users/lenovo/Downloads/employe data.csv"))
projects=data.frame(project)
projects
```

#Dimension of the dataset
```{r}
dim(projects)
```
#Head()/retrive the first six rows
```{r}
head(projects)
```
#summary()/summarise the data
```{r}
summary(projects)
```

#na.omit functions to used to remove the missing values
```{r}
projects=na.omit(projects)
projects
```
#colsums is used to check whether the missing values or not 
```{r}
colSums(is.na(projects))
```
#no of variables in the employee dataset along no of rows & columns using cat() function to show that.
```{r}
cat("No. of Rows:",nrow(projects),"\nNo. of Columns:",ncol(projects),"\n\n")
str(project)
```

#Random Number Generator (Reproduce pupose)
```{r}
set.seed(42)
```

#splitting the dataset.# train_data & test_data
```{r}
train_indices = sample(1:nrow(projects),size=0.8*nrow(projects)) #MODEL SPLITTING
train_data = projects[train_indices,]
test_data = projects[-train_indices,]
```

#After splitted the dataset 
```{r}
cat(
  "No. of Rows (in Test Dataset):",nrow(projects),
  "\nNo. of Columns (in Test Dataset):",ncol(test_data),
  "\n\nNo. of Rows (in Training Dataset):",nrow(train_data),
  "\nNo. of Columns (in Training Dataset):",ncol(train_data)
    )
```

```{r}
train_data
```

```{r}
test_data
```

#linear model regression/test_data
```{r}
model=lm(satisfaction_level~Emp.ID+salary+promotion_last_5years+average_montly_hours+last_evaluation+time_spend_company+number_project,data=test_data)
summary(model)
```
#linear regression model/train_data
```{r}
model=lm(satisfaction_level~Emp.ID+salary+promotion_last_5years+average_montly_hours+last_evaluation+time_spend_company+Work_accident+number_project,data=train_data)
summary(model)
```

#this plot clearly provide the insight into how satisfaction level are distributed across employees.
```{r}
ggplot(projects,aes(x=satisfaction_level))+geom_histogram(bins=30,fill="blue",color="black")+ggtitle("distribution of employee satisfaction levels")+xlab("satisfaction level")+ylab("frequency")+theme_minimal()
```
#this plot show the distribution of satisfaction levels for employees in each department,including median,interquartile range,and potential outliers

```{r}
ggplot(train_data,aes(x=dept,y=satisfaction_level,fill=dept))+geom_boxplot()+ggtitle("satisfaction levels across department")+xlab("department")+ylab("satisfaction levels")+theme_minimal()+theme(axis.text.x = element_text(angle = 45,hjust = 1))+scale_fill_brewer(palette = "set3")
```
#this plot shows how salary levels are distributed among employees based on their promotion status in the last 5 years.
```{r}
ggplot(data = projects,aes(x=promotion_last_5years,fill=salary))+geom_bar(position = "dodge")+ggtitle("promotion_last_5years vs salary")+xlab("promotion last 5 years")+ylab("count")+scale_fill_brewer(palette = "set2",name="salary")+theme_minimal()
```
```{r}
ggplot(projects,aes(x=Emp.ID,y=satisfaction_level,color=salary))+geom_point(size=3,alpha=0.7)+scale_color_gradient(low="white",high="black")+labs(title = "scatter plot of Emp.ID vs satisfaction level by salary",x="Emp.ID",y="satisfaction level",color="salary")+theme_minimal()
```


